Technical Report

### 2018-06-18 fully-abstract-comp/technical-report.tex
first stap at global ret and clos seals.

\[
  \lrvtrusted[\square,(\ta,\stkb)](W) =
  \begin{array}[t]{l}
    \lrv(W)\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
    \; \middle| \;
    \begin{array}{l}
      \exists r \in \dom(\pwheap) \ldotp \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)} \wedge\\
      \hfill\dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss)
    \end{array}
    \right\} \cup \\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\normal),\baddr,\eaddr,\aaddr)},\\ ((\perm,\normal),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{l}
      \perm \sqsubseteq \rx \wedge \\
      \npair{[\baddr,\eaddr]} \in \xReadCond[\square,(\ta,\stkb)]{W}
    \end{array}
    \right\}
  \end{array}
\]

@@

\begin{equation*}
  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_) \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
    \begin{array}{l}
    \dom(\code) = [\aaddr,\baddr] \wedge \\
      (\dom(\code) \subseteq \ta \wedge \trust = \trusted) \vee \\
      \quad (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
      \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \wedge\\
      \forall a \in \dom(\code), n' < n \ldotp\\
      \quad\npair[n']{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
    \end{array}
  \right\}
\end{equation*}

\lau{19-03-2018: Add the link capability to the above? code in ta?}
\lau{TODO: look at the $[b,e] \subseteq ...$ condition - will it work for splice?}
\begin{multline*}
  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
  (\ta,\stkb) \; \sigma \; \hat{W}= \\
  \begin{array}[t]{l}
\left\{
    \begin{array}{l}
\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrc(\baddr,\eaddr,\aaddr'+\calllen),\\((\rx,\normal),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
      \begin{array}{l}
        \dom(\code) \subseteq \ta \tand\\
        \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
        \aaddr = \aaddr' + \retoffset \tand \\
        \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
        \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
      \end{array}
    \end{array}
      \right\} \uplus \\
\left\{
    \begin{array}{l}
\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrd(\baddr,\eaddr),\\((\rw,\linear),\baddr,\eaddr,\baddr-1)\endarray\right)} \middle| \right. \\
      \begin{array}{l}
        \dom(\code) \subseteq \ta \tand\\
        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
        \quad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
        \quad \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
        \quad \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
      \end{array}
    \end{array}
    \right\} \\
  \end{array}\\
    \text{ if } \sigma \in \sigrets\\
\end{multline*}
and
\begin{multline*}
  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
  (\ta,\stkb) \; \sigma \; \hat{W}= \\
  \left\{
    \begin{array}{l}
\left. \npair{(\vsc, \vsc' )} \middle| \right. \\
      \begin{array}{l}
        \sigma \in \sigcloss \tand\\
        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
        (\dom(\code) \subseteq \ta \tand \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W}))
      \end{array}
    \end{array}
  \right\}\\
  \text{ if } \sigma \in \sigcloss
\end{multline*}

@@

\dominique{The below is an experimental reformulation of the reasonability requirement for the FTLR and full abstraction formulation.  All remarks welcome..
  The main change is that it now says that after a call, the post-return code is also expected to behave reasonably under reasonable assumptions. 
}
\dominique{TODO: better name than ``reasonable'' for all of this.}
Take a block of code memory $\mscode$, a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\sigrets$ and $\sigcloss$.
We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ if $n=0$
or the following implications hold.
\begin{itemize}
\item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\sigrets \cup \sigcloss)$
\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$ and $\sigrets \neq \emptyset$, then $[\baddr,\eaddr] \# \dom(\mscode)$
\item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\sigrets\cup \sigcloss)$ then $\vsc$ is reasonable up to $n - 1$ steps.
\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
\end{itemize}

We say that an execution configuration $\Phi$ is reasonable up to $n$ steps if:
  \begin{itemize}
  \item \emph{Guarantee stack base address before call} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
    \item $\Phi'$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
    \item $0 < i \le n$
    \end{itemize}
    Then all of the following hold:
    \begin{itemize}
    \item $\src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\eaddr_\stk,\aaddr_\stk}}$
    \item $r_1 \neq \rtmp{1}$
    \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behaves reasonably up to $n-i$ steps
    \end{itemize}
  \item \emph{Use return seals only for calls} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
    \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
    \item $0 < i \le n$
    \end{itemize}
    Then one of the following holds:
    \begin{itemize}
    \item $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \sigrets$ 
    \item $\sigma \in \sigcloss$.
    \end{itemize}
  \item \emph{Don't store private stuff...} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
    \item $\src{\Phi'}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then 
    \item $0 < i \le n$
    \end{itemize}
    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps
  \item \emph{Don't leak private stuff...} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi' \step \Phi''$
    \item $\src{\Phi'}$ is in $\ta$
    \item $\src{\Phi''}$ is not in $\ta$
    \item $0 < i \le n$
    \end{itemize}
    Then $\Phi.\reg(r)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps for all $r$ (including $\pcreg$)
  \end{itemize}

>>>>

\[
  \lrvtrusted[\square,(\ta,\stkb,\gsigrets,\gsigcloss)](W) =
  \begin{array}[t]{l}
    \lrv(W)\cup \\
%
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{\seal{\sigma,\sigma_\baddr,\sigma_\eaddr}},\\ \seal{\sigma,\sigma_\baddr,\sigma_\eaddr} \endarray \right)} 
    \; \middle| \;
    \begin{array}{l}
      \exists r \in \dom(\pwheap) \ldotp \\
      \quad \pwheap(r) \nequal \codereg{\sigrets,\sigcloss,\code,(\ta,\stkb)} \wedge\\
      \quad \dom(\code) \subseteq \ta \wedge [\sigma_\baddr,\sigma_\eaddr] \subseteq (\sigrets\cup\sigcloss) \wedge\\
      \quad \sigrets \subseteq \gsigrets \wedge \sigcloss \subseteq \gsigcloss
    \end{array}
    \right\} \cup \\
    \left\{ \npair{\left(\arraycolsep=0pt\array{l} \src{((\perm,\normal),\baddr,\eaddr,\aaddr)},\\ ((\perm,\normal),\baddr,\eaddr,\aaddr) \endarray \right)} \;\middle|\; 
    \begin{array}{l}
      \perm \sqsubseteq \rx \wedge \\
      \gc = (\ta,\stkb,\gsigrets,\gsigcloss)  \wedge \\
      \npair{[\baddr,\eaddr]} \in \xReadCond[\square,\gc]{W} 
    \end{array}
    \right\}
  \end{array}
\]

@@

\begin{equation*}
  H^\mathrm{code} \; \sigrets \; \sigcloss \; \code \; (\ta,\_,\gsigrets,\gsigcloss) \; \hat{W} = \left\{\npair{(\code,\code)} \middle|
    \begin{array}{l}
    \dom(\code) = [\aaddr,\baddr] \wedge \\
      (\dom(\code) \subseteq \ta \wedge \sigrets \subseteq \gsigrets \trust = \trusted) \vee \\
      \quad (\dom(\code)\mathrel{\#} \ta \wedge \sigrets = \emptyset \wedge \trust =\untrusted) \wedge \\
      \sigrets,\sigcloss \vdash_{\mathrm{comp-code}} \code \wedge\\
      \forall a \in \dom(\code), n' < n \ldotp\\
      \quad\npair[n']{(\code(a),\code(a))} \in \lrvg{\trust}(\purePart{\xi(\hat{W})})
    \end{array}
  \right\}
\end{equation*}

\lau{19-03-2018: Add the link capability to the above? code in ta?}
\lau{TODO: look at the $[b,e] \subseteq ...$ condition - will it work for splice?}
\begin{multline*}
  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
  (\ta,\stkb) \; \sigma \; \hat{W}= \\
  \begin{array}[t]{l}
\left\{
    \begin{array}{l}
\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrc(\baddr,\eaddr,\aaddr'+\calllen),\\((\rx,\normal),\baddr,\eaddr,\aaddr)\endarray\right)} \middle| \right. \\
      \begin{array}{l}
        \sigrets \subseteq \gsigrets \tand \\
        \dom(\code) \subseteq \ta \tand\\
        \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
        \aaddr = \aaddr' + \retoffset \tand \\
        \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets \tand\\
        \lbrack \aaddr',\aaddr' + \calllen -1 \rbrack \subseteq \lbrack \baddr, \eaddr \rbrack
      \end{array}
    \end{array}
      \right\} \uplus \\
\left\{
    \begin{array}{l}
\left. \npair{\arraycolsep=0pt\left(\array{l}\retptrd(\baddr,\eaddr),\\((\rw,\linear),\baddr,\eaddr,\baddr-1)\endarray\right)} \middle| \right. \\
      \begin{array}{l}
        \sigrets \subseteq \gsigrets \tand \\
        \dom(\code) \subseteq \ta \tand\\
        \exists r \in \addressable{\linear,\pwpriv[\xi(\hat{W})]} \ldotp \pwpriv[\xi(\hat{W})](r).H \nequal (\stareg[(\ms_S,\ms_T),(\ta,\stkb)]{\spao,\square}, \aaddr'+\calllen) \tand \\
        \quad \dom(\ms_S) = \dom(\ms_T) = [\baddr,\eaddr] \tand\\
        \quad \decInstr{\code([\aaddr',\aaddr' + \calllen-1])} = \overline{\scall{\offpc,\offsigma}{r_1}{r_2}} \tand \\
        \quad \code(\aaddr'+\offpc) = \seal{\sigma_b,\sigma_e,\sigma_b} \tand \sigma = \sigma_b + \offsigma \in \sigrets
      \end{array}
    \end{array}
    \right\} \\
  \end{array}\\
    \text{ if } \sigma \in \sigrets\\
\end{multline*}
and
\begin{multline*}
  H^\mathrm{code,\square}_\sigma \; \sigrets \; \sigcloss \; \code \;
  (\ta,\stkb) \; \sigma \; \hat{W}= \\
  \left\{
    \begin{array}{l}
\left. \npair{(\vsc, \vsc' )} \middle| \right. \\
      \begin{array}{l}
        \sigma \in \sigcloss \tand\\
        (\dom(\code) \mathrel{\#} \ta \tand \npair{(\vsc,\vsc')} \in \lrv \; \xi(\hat{W})) \tor\\
        (\dom(\code) \subseteq \ta \tand \sigcloss \subseteq \gsigcloss \npair{(\vsc,\vsc')} \in \lrvtrusted \; \xi(\hat{W}))
      \end{array}
    \end{array}
  \right\}\\
  \text{ if } \sigma \in \sigcloss
\end{multline*}

@@

\dominique{The below is an experimental reformulation of the reasonability requirement for the FTLR and full abstraction formulation.  All remarks welcome..
  The main change is that it now says that after a call, the post-return code is also expected to behave reasonably under reasonable assumptions. 
}
\dominique{TODO: better name than ``reasonable'' for all of this.}
Take a block of code memory $\mscode$, a set of trusted addresses $\ta$ and sets of return pointer and closure seals $\gsigrets$ and $\sigcloss$.
We define that a word $w$ is reasonable up to $n$ steps in memory $\ms$ if $n=0$
or the following implications hold.
\begin{itemize}
\item If $w = \seal{\sigma_\baddr,\sigma_\eaddr,\_}$, then $[\sigma_\baddr,\sigma_\eaddr] \mathrel{\#} (\gsigrets \cup \gsigcloss)$
\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \neq 0$ and $\gsigrets \neq \emptyset$, then $[\baddr,\eaddr] \# \dom(\mscode)$
\item If $w = \sealed{\sigma,\vsc}$ and $\sigma \not\in (\gsigrets\cup \gsigcloss)$ then $\vsc$ is reasonable up to $n - 1$ steps.
\item If $w = ((\perm,\_),\baddr,\eaddr,\_)$ and $\perm \in \readAllowed{}$ and $n > 0$, then $\ms(\aaddr)$ is reasonable up to $n - 1$ steps for all $\aaddr \in ([\baddr,\eaddr] \setminus \ta)$
\end{itemize}

We say that an execution configuration $\Phi$ is reasonable up to $n$ steps if:
  \begin{itemize}
  \item \emph{Guarantee stack base address before call} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
    \item $\Phi'$ points to $\src{\scall{\offpc,\offsigma}{r_1}{r_2}}$ in $\ta$ for some $\src{r_1}$ and $\src{r_2}$
    \item $0 < i \le n$
    \end{itemize}
    Then all of the following hold:
    \begin{itemize}
    \item $\src{\Phi'}(\src{r_\stk}) = \src{\stkptr{\_,\stkb,\eaddr_\stk,\aaddr_\stk}}$
    \item $r_1 \neq \rtmp{1}$
    \item $\Phi'(\pcreg) + \calllen$ and $\Phi'(\rstk)$ behaves reasonably up to $n-i$ steps
    \end{itemize}
  \item \emph{Use return seals only for calls} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
    \item $\src{\Phi'}$ points to $\src{\tcseal{r_1}{r_2}}$ in $\ta$ and $\Phi'(r_1) = \seal{\sigma_\baddr,\sigma_\eaddr,\sigma}$
    \item $0 < i \le n$
    \end{itemize}
    Then one of the following holds:
    \begin{itemize}
    \item $\src{\Phi'}$ is inside $\src{\scall{\offpc,\offsigma}{r_1'}{r_2'}}$ and $\sigma \in \gsigrets$ 
    \item $\sigma \in \gsigcloss$.
    \end{itemize}
  \item \emph{Don't store private stuff...} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi'$
    \item $\src{\Phi'}$ points to $\src{\tstore{r_1}{r_2}}$ in $\ta$, then 
    \item $0 < i \le n$
    \end{itemize}
    Then $\Phi.\reg(r_2)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps
  \item \emph{Don't leak private stuff...} If
    \begin{itemize}
    \item $\Phi \nstep[i]{\ta,\stkb} \Phi' \step \Phi''$
    \item $\src{\Phi'}$ is in $\ta$
    \item $\src{\Phi''}$ is not in $\ta$
    \item $0 < i \le n$
    \end{itemize}
    Then $\Phi.\reg(r)$ is reasonable in memory $\Phi'.\mem$ up to $n$ steps for all $r$ (including $\pcreg$)
  \end{itemize}
